Balish Practical Machine Learning
========================================================
Background per assignment:

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively.  One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

Data provided per web link includes a training and a test set. I will read in the data (training) and the needed caret library and explore some features:

```{r}
library(caret)
pmltrain<-read.csv("pml-training.csv")
summary(pmltrain)

```

Goal:
------------------------------------------------------------------
The goal of the project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did.



```{r}
# There are many variables with mainly NA and use the following code to eliminate

pmltrainok<-pmltrain[,!apply(is.na(pmltrain), 2, any)]

#There remain variables with near zero variance and some of which have undefined
#values. Will find these and then also remove them leaving the pmltrainred dataset

nsv <- nearZeroVar(pmltrainok,saveMetrics=TRUE)
good<-rownames(nsv)[nsv$nzv==FALSE]
pmltrainred<-pmltrainok[,names(pmltrainok)%in%good]

```
The first 6 variables are time and subject names and numbers and will also eliminate thosewith the following:

```{r}
pmltrainokk<-pmltrainred[,-c(1:6)]

#Will set seed and then create a training and testing set from the original training set to allow some validation.

set.seed(3455)
inTrain <- createDataPartition(y=pmltrainokk$classe,
                                 p=0.4, list=FALSE)
training <- pmltrainokk[inTrain,]
testing <- pmltrainokk[-inTrain,]
```
The classe variable is now the 53rd column and will look at the degree of correlations among variables:


```{r}
M <- abs(cor(training[,-53]))
diag(M) <- 0
which(M > 0.8,arr.ind=T)

```
```{r fig.width=7, fig.height=6}
plot(cars)
```

